---
title: "Aides Drug"
author: "Alexis Velasco"
date: "6/14/22"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
triUmp <-read.csv(file="triUmpsal.csv",header=TRUE)
```

```{r eval=FALSE}
triUmp <- triUmp %>%
mutate(across(where(is.character), as.factor))
```

```{r}
females<- filter(triUmp,gender=="F")
```



# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

A new  preventative aides drug (triUmpsal) is under test in clinical trials. Consider only the females, does this drug seem effective?  Hint - use dplyr to obtain that subpopulation with the filter command.

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->

```{r}
barchartGC(~infected + treatment,data=females)
barchartGC(~infected + treatment,data=females, type="percent")

```

Based on the graphs results females who took the placebo seem to have a smaller chance to not get AIDS, but those who had taken the drug still contracted AIDS rather than those who took the placebo drug.

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
table1 <- xtabs(~infected + treatment,data=females)
rowPerc(table1)
colPerc(table1)
```

The numerical results show slightly over 50% of females that took the AIDS drug still got AIDS, where 41% of females who took the placebo. The results show females who took the drug to prevent AIDS were more likely to get the virus, showing the drug doesn't not work properly. 

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The inferential results shows 218 people who took the drug still contracted AIDS where only 147 who took the placebo got AIDS. There were more who took the drug that contracted AIDS than those who took the placebo. The chances of getting these results are 7 out of 10,000 if there was no diffrence , a number so small leads to the belief that it was not by chance those who took the drug got AIDS more than those who took the placebo. The drug seems to be working in favor of contracting AIDS rather than helping the prevention of AIDS, allowing us to reject the null hypothesis. 

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->


Based on the graphs results females who took the placebo seem to have a smaller chance to not get AIDS.The numerical results show slightly over 50% of females that took the AIDS drug still got AIDS, where 41% of females who took the placebo.The inferential results shows 218 females who took the drug still contracted AIDS where only 147 who took the placebo got AIDS. There were more females who took the drug that contracted AIDS than those who took the placebo. The Fisher exact test shows that those who take the drug are 1.240% between 1.92% more likely to contract AIDS. Having these results leads us to rejecting the null hypothesis because the drug in working in favor of contracting AIDS rather that preventing AIDS. The drug to help prevent AIDS and contracting AIDS shared a relationship by allowing those who took it to have a higher chance of contracting AIDS than those who took the placebo. 

